---
title: "EDA Assignment2"
author: "Motoki Sugino"
date: "August 30, 2016"
output:
  word_document: default
  html_document: default
---
#PART A:  Working with a Single Batch – Displays
```{r}
##1.  The dataset baseball.attendance in the LearnEDA package contains the average (mean) home attendance of the 30 Major League baseball teams in 2010.  For this exercise, you don’t have to understand baseball.  All you have to know is that there are 30 baseball teams and the names of the teams correspond to their city locations.  There are two leagues – the National League has 16 teams and the American League has 14 teams.  The dataset has four columns:

#Team – the name of the team (city)
#League – the league that this team belongs
#N.Home – the number of home games
#Avg.Home – the mean attendance (number of people) at games played at home
#Pct.Home – the percentage of park capacity for home games
#N.Away – the number of away games
#Avg.Away – the mean attendance (number of people) at games played away from home
#Pct.Away – the percentage of park capacity for away games

#For this assignment, we focus on the average home attendance (variable Avg.Home).
library(LearnEDA)
#baseball.attendance
```
(a)  Construct a stemplot where

-        the breakpoint is between the 10,000 and 1000 places
-        you have 10 leaves per stem

 Discuss what you see in this distribution of attendance numbers, including shape, “average”, spread, and any unusual characteristics.  If there are any unusual high or low attendances, identify the cities that have these unusual mean attendances.
```{r}
baseball.attendance[,c("Team","Avg.Home")]
aplpack::stem.leaf(baseball.attendance$Avg.Home,m=1, depth=FALSE)#10 leaves per stem
```

*Looking at a data distribution, to me it is slightly skewed right. Most of the attendance average fall in the 2-3 range and the values at high end decrease at a much slower than the values at the low end. 
Cleveland has the lowest average attendance. NY_Yankees has the highest average attendance.
I do not think that there are unusual characteristics at this time. However, NY_Yankees, Philadelphia and LA_Dodgers might be unusual high attendances. Florida and Cleveland might be unusual low attendances. 

(b)  Redraw the stemplot using the same breakpoint and 5 leaves per stem.
```{r}
aplpack::stem.leaf(baseball.attendance$Avg.Home,m=2, depth=TRUE)#5 leaves per stem
```


(c)  Redraw the stemplot using the same breakpoint and 2 leaves per stem.
```{r}
aplpack::stem.leaf(baseball.attendance$Avg.Home,m=5, depth=FALSE)#2 leaves per stem
```

(d)  Discuss any features that you’ve learned about this dataset by constructing these two additional stemplots.  What is the best choice of stemplot?  Why?

*The 2-leaves-per-stem stemplot looks pretty good to me because you can see three clusters in this stemplot. Therefore, I chose the 2-leaves-per-stem stemplot as the best one in this dataset.

2.   In Assignment 1, you found two datasets. 

For each dataset …

(a) Read the data into R.

```{r}
Pokemondata <- read.csv("https://docs.google.com/spreadsheets/d/1ni9fGCOfnUd3BYuggKm_jttRh8DHpQFgLXHiZYDPiCg/pub?output=csv")
SalariesForSF <- read.csv("https://docs.google.com/spreadsheets/d/1tqh_FERElW3C0Z_Rn2NP76SZdsbKsrViqjw_vtQEFww/pub?output=csv")
```

(b)   Construct two different stemplots using R.

```{r}
aplpack::stem.leaf(Pokemondata$Total, m=5, depth=FALSE)#2 leaves per stem
aplpack::stem.leaf(Pokemondata$Total, m=1, depth=FALSE)#10 leaves per stem

aplpack::stem.leaf(SalariesForSF$Regular.pay, m=5, depth=FALSE)#2 leaves per stem
aplpack::stem.leaf(SalariesForSF$Regular.pay, m=1, depth=FALSE)#10 leaves per stem
```


(c)  Discuss the data (shape, average, spread, unusual characteristics, etc.) and discuss the best choice of stemplot.

*For Pokemondata, it is the almost symmetric to me. The majority of the data is in the middle and the value drops off at the same rate at the low end and right end. The large number of total averages in 4 line, so 400 something would be a typical value. The most of the total averages fall 300 and 500, with only a couple of pokemon with averages below 200 or above 500. Some pokemon have total above 770. They might be unusually high values. In addition, some pokemon have total below 200. They might be unusually low values. However, I would say that there are no unusual values in the Polemondata.
To me the best choice of stemplot is 2-leaves-per-stem.


*For SalariesForSF, it is the almost symmetric if you see the 10-leaves-per-stem stemplot. However, there are two bumps in the data distribution if you look at 2-leaves-per-stem stemplot. The the large number of regular pay averages in 2 if you look at 10-leaves-per-stem stemplot. The most of the regular pay averages fall  130000 and 290000, with only a couple of regular pay with averages below 100000 or above 300000. There is one unusually high value which is 507831.6.
To me the best choice of stemplot is 2-leaves-per-stem.


#PART B:  Working with a Single Batch – Summaries

 For each of the three datasets (the baseball attendance data and the two interesting datasets that you found)

(a)  compute the letter values (median, fourths, eighths, and extremes)
```{r}
####For baseball 
#depth of median
n=30
(n+1)/2

#depth of fouth
(15+1)/2

#depth of eight
(8+1)/2

#extremes
fivenum(baseball.attendance$Avg.Home)
45715.0
####For pokemon dataset

#depth of median
n=800
(n+1)/2

#depth of fouth
(400+1)/2

#depth of eight
(200+1)/2

#extremes
fivenum(Pokemondata$Total)
780
####For SalariesForSF

#depth of median
n=50
(n+1)/2

#depth of fouth
(25+1)/2

#depth of eight
(13+1)/2

#extremes
fivenum(SalariesForSF$Regular.pay)
507831.60
```


(b)    find the mean and median.  If these two measures of average are different, explain why (look at the shape of the dataset).
```{r}
#median and mean for baseball dataset
baseball.median <- median(baseball.attendance$Avg.Home);baseball.median
baseball.mean <- mean(baseball.attendance$Avg.Home);baseball.mean

#median and mean for pokemon dataset
pokemon.median <- median(Pokemondata$Total);pokemon.median
pokemon.mean <- mean(Pokemondata$Total);pokemon.mean

#median and mean for SalariesForSF
SalariesForSF.median <- median(SalariesForSF$Regular.pay);SalariesForSF.median
SalariesForSF.mean <- mean(SalariesForSF$Regular.pay);SalariesForSF.mean
```

(c)    find the fourth spread
```{r}
####For baseball 
fivenum(baseball.attendance$Avg.Home)
dF=37592.0-22850.0;dF
####For pokemon dataset
fivenum(Pokemondata$Total)
dF=515-330;dF
####For SalariesForSF
fivenum(SalariesForSF$Regular.pay)
dF=279311.03-180314.37;dF

```

(d)    find the step, the inner and outer lower and upper fences, and identify any outliers (mild and extreme) in your dataset.  Can you offer any explanation for these outliers?
```{r}
####For baseball 
STEP=1.5*14742;STEP
FenceL=22850.0-STEP;FenceL
FenceU=37592.0+STEP;FenceU
FENCEL=22850.0-(2*STEP);FENCEL
FENCEU=37592.0+(2*STEP);FENCEU

#Do we have any outliers? No

####For pokemon dataset
STEP=1.5*185;STEP
FenceL=330-STEP;FenceL
FenceU=515+STEP;FenceU
FENCEL=330-(2*STEP);FENCEL
FENCEU=515+(2*STEP);FENCEU

#Do we have any outliers? No

####For SalariesForSF
STEP=1.5*98996.66;STEP
FenceL=180314.37-STEP;FenceL
FenceU=279311.03+STEP;FenceU
FENCEL=180314.37-(2*STEP);FENCEL
FENCEU=279311.03+(2*STEP);FENCEU

#Do we have any outliers? YES,there is one outlier. it makes because he is a chief Investment officer in San Francisco. He makes a lot of money more than others. Some of people should be reich in San Francisco in capitalism.
```

