---
title: "EDA Assignment 4"
author: "Motoki Sugino"
date: "September 29, 2016"
output: html_document
---

1.  The dataset lake in the LearnEDA package  (taken from the Minitab dataset collection) contains measurements of lakes in the Vilas andOneidacounties of northernWisconsin.  
The variables are

AREA = area of lake in acres

DEPTH = maximum depth of lake in feet

PH = pH (acidity) measurement

WSHED = watershed area in square miles

HIONS = concentration of hydrogen ions

For two of these variables (that are not symmetric), find a transformation which makes the dataset symmetric. You can use any tool you want, including H quick method, inspection of mids, symmetry plots, etc. Demonstrate that your transformation has been symmetric.

```{r}
library(LearnEDA)
#lake
```

looking at each histgram
```{r}
par(mfrow=c(2, 2))
hist(lake$Area)
#hist(lake$Depth)
#hist(lake$PH)
hist(lake$Wshed)
hist(lake$Hion)
```

The variables, Area,Wshed and Hion, are not symmetric. Therefore, we are going to choose the variable Area and Wshed to perfrom the transformation for symmetric distribution.

#For Area
Now we are going to focus on the variable Area.

Shape of data
```{r}
aplpack::stem.leaf(lake$Area)
hist(lake$Area, freq=FALSE)
lines(density(lake$Area, bw=150),lwd=2)
```
There is rigth skewedness in the distribution.

The power family 
```{r}
par(mfrow=c(2, 2))
hist(lake$Area, main="RAW")
hist(sqrt(lake$Area), main="ROOTS")
hist((lake$Area) ^ 0.001, main="P = 0.001")
```

From the above plots, we can say that taking p=0.001 looks rougly symmetric.

Cheacking for symmetry by looking at midsummaries.
```{r}
letter.values <- lval(lake$Area)
letter.values$mids
plot(letter.values$mids)
```

You can see the right skewness since this sequence is in creasing.

We first try roots (p = .5).
```{r}

roots <- sqrt(lake$Area)
aplpack::stem.leaf(roots)
root.lv <- lval(roots)
root.lv
plot(root.lv$mids)
```

Next we try log (p=.5).
```{r}
logs <- log(lake$Area) 
aplpack::stem.leaf(logs)
logs.lv <- lval(logs)
logs.lv
plot(logs.lv$mids)

```

Looking the above graphs, we can see the taking log looks like the best reexpression.
We are going to take a further investigation to check if taking log is the best. 
```{r}
#if take p = −1/2     -1/sqrt(lake$Area) Further investigation 
recroots <- - 1 / sqrt(lake$Area)
aplpack::stem.leaf(recroots)
recroots.lv <- lval(recroots)
recroots.lv
plot(recroots.lv$mids)
```

Once again, taking log is the best option for this case since stemplot looks like more symmetric.

Next, we use Hinkley's quick method.
```{r}
hinkley(roots)
hinkley(logs)
hinkley(recroots)
```

Looking at the values of d in Hinkley's quick method, the correct reexpression appears to be between p=.5 and p=0, although the value closest to 0 corresponds to the log reexpression. 

Checking at Symmetry Plot.
```{r}
symplot(lake$Area)
symplot(logs)
```

For Area, taking log is the best reexpression for Area.Looking at the symmetry plot, the points roughly fall close to the line u=v, so we can say that the data is neraly symmetric.
However, there is right skewedness in the end and one possible outlier. 

Matched transformation.
```{r}
#let x_0 = 148, the medain of the raw data
matched.roots <- 148 + (sqrt(lake$Area) - sqrt(148)) / (.5*148 ^ (-.5))
matched.logs <- 148 + (log10(lake$Area) - log10(148)) / (log10(exp(1)) / 148)

raw <-lake$Area
matched.roots <- mtrans(raw, 0.5)
matched.logs <- mtrans(raw, 0)
boxplot(data.frame(raw, matched.roots, matched.logs))
```

Looking at the boxplots, we can see that all of data look like symmetric.
However, while raw date has wide spread of data distribution including outliers, matched.log has the narrow data spread.

In conclusion, taking log is the best reexpression for the variable, Area, since it removes right skewedness and it looks like approximately symmetric. However, there is right skewedness in the end and one possible outlier as you can see it in the symmetry plot. 

#For Wshed
```{r}
aplpack::stem.leaf(lake$Wshed)
hist(lake$Wshed, freq=FALSE)
lines(density(lake$Wshed, bw=4),lwd=2)
```

Once again, there is right skewedness in the distribution.

```{r}
par(mfrow=c(2, 2))
hist(lake$Wshed, main="RAW")
hist(sqrt(lake$Wshed), main="ROOTS")
hist((lake$Wshed) ^ 0.001, main="P = 0.001")
```

These histgram above have no symmetric. 

```{r}
letter.values <- lval(lake$Wshed)
letter.values$mids
plot(letter.values$mids)
```

You can see the right skewness since this sequence is in creasing.

taking roots
```{r}
roots <- sqrt(lake$Wshed)
aplpack::stem.leaf(roots)
root.lv <- lval(roots)
root.lv
plot(root.lv$mids)
```


```{r}
#take log
logs <- log(lake$Wshed) 
aplpack::stem.leaf(logs)
logs.lv <- lval(logs)
logs.lv
plot(logs.lv$mids)
```

There is still right skewedness in the tail here.
Taking log does not looks roughly symmetric.

```{r}
#if take p = −1/2     -1/sqrt(lake$Area) Further investigation 
recroots <- - 1 / sqrt(lake$Wshed)
aplpack::stem.leaf(recroots)
recroots.lv <- lval(recroots)
recroots.lv
plot(recroots.lv$mids)

hinkley(roots)
hinkley(logs)
hinkley(recroots)
```

This plot above looks better than the plot with taking log. 
  
Looking at the values of d in Hinkley's quick method, the correct reexpression appears to be between p=0 and p=-1/2, although the value closest to 0 corresponds to the log reexpression. 

Now, we are going to take an alternative form of this power transformation that is p=-0.2
```{r}
#An Alternative form of this power transformation
new <-((lake$Wshed ^ -0.2)-1)/-0.2
aplpack::stem.leaf(new)
hist(((lake$Wshed ^ -0.2)-1)/-0.2)
```

This histgram looks roughly symmetric.

```{r}
new.lv <- lval(new)
new.lv
plot(new.lv$mids)
#Symmetry Plot
symplot(new)
```

We have gotten the best reexpression which is ((lake$Wshed ^ -0.2)-1)/-0.2. Looking at the symmetry plot, the points roughly fall close to the line u=v, so we can say that the data is neraly symmetric. However, there is one outlier in the end.

```{r}
#let x_0 = 0.389, the medain of the raw data
matched.p <- 0.389 + (((lake$Wshed ^ -0.2)-1)/-0.2 - ((lake$Wshed ^ -0.389)-1)/-0.389 / lake$Wshed ^ -0.8)
matched.logs <- 0.389 + (log10(lake$Wshed) - log10(0.389)) / (log10(exp(1)) / 0.389)

raw <-lake$Wshed
matched.p <- mtrans(raw, 0.5)
matched.logs <- mtrans(raw, 0)
boxplot(data.frame(raw, matched.p, matched.logs),ylim=c(-5,20))

par(mfrow=c(2, 1))
boxplot(((lake$Wshed ^ -0.2)-1)/-0.2, horizontal = TRUE)
boxplot(lake$Wshed, horizontal = TRUE,ylim=c(0,20))
```

From the boxplot for ((lake$Wshed ^ -0.2)-1)/-0.2, we can see the data have symmetric shape now. 
In conclusion, the best reexpression which is ((lake$Wshed ^ -0.2)-1)/-0.2 since it gives the approximately symmetric shape of data spreads. Moreover, some of the outliers are removed by this transformation. 

2.  Find two datasets (at least 30 observations in each) that you are interested in that are not symmetric.  Find a transformation that makes each dataset symmetric and demonstrate that your transformation is effective in achieving approximate symmetry.

#For Pokemondata
We are going to check the data spread of HP data in pokemondata set.
```{r}
Pokemondata <- read.csv("https://docs.google.com/spreadsheets/d/1ni9fGCOfnUd3BYuggKm_jttRh8DHpQFgLXHiZYDPiCg/pub?output=csv")
```

```{r}
hist(Pokemondata$HP)
par(mfrow=c(1, 1))
hist(sqrt(Pokemondata$HP + 0.5), main="ROOTS")
lval(Pokemondata$HP)
```

There is right skewedness in the data.

```{r}
letter.values <- lval(Pokemondata$HP)
letter.values$mids
plot(letter.values$mids)
```

You can see the right skewness since this sequence is in creasing. 

```{r}
#take roots
roots <- sqrt(Pokemondata$HP)
aplpack::stem.leaf(roots)
root.lv <- lval(roots)
root.lv
plot(root.lv$mids)

#take log
logs <- log(Pokemondata$HP) 
aplpack::stem.leaf(logs)
logs.lv <- lval(logs)
logs.lv
plot(logs.lv$mids)

```

Taking roots looks like the best reexpression.

```{r}
#if take p = −1/2     -1/sqrt(lake$Area) Further investigation 
recroots <- - 1 / sqrt(Pokemondata$HP)
aplpack::stem.leaf(recroots)
recroots.lv <- lval(recroots)
recroots.lv
plot(recroots.lv$mids)

hinkley(roots)
hinkley(logs)
hinkley(recroots)
```

Taking roots is the best reexpression to me. However, there are many outliers next to both side of whiskers. Looking at the values of d in Hinkley's quick method, the correct reexpression appears to be between p=.5 and p=0, although the value closest to 0 corresponds to the log reexpression.


Therefore, we are going to take a new power transforation as p=0.2
```{r}
new<-(Pokemondata$HP) ^ 0.2
hist((Pokemondata$HP) ^ 0.2, main="P = 0.2")
aplpack::stem.leaf((Pokemondata$HP)^0.2)
new.lv <- lval((Pokemondata$HP) ^ 0.2)
new.lv
plot(new.lv$mids)
symplot(new)
boxplot(new)
```

Looking at a symmetry plot with p=0.2, you can see the data was reexpressed as roughly symmetric.
Therefore, (Pokemondata$HP) ^ 0.2 is the best transformation.

```{r}
#let x_0 = 65, the medain of the raw data
x_0<-65
matched.roots <- x_0 + (sqrt(Pokemondata$HP) - sqrt(x_0)) / (.5*x_0 ^ (-.5))
matched.logs <- x_0 + (log10(Pokemondata$HP) - log10(x_0)) / (log10(exp(1)) / x_0)

raw <-Pokemondata$HP
matched.roots <- mtrans(raw, 0.5)
matched.logs <- mtrans(raw, 0)
boxplot(data.frame(raw, matched.roots, matched.logs))
```

Note that this matching has given the three batches the same median (x_0) and the batches have similar spreads. 
All of data look like symmetric. While raw data has one outlier next to the left whisker, matched.logs data has three outliers next to the left whisker. 

In conclusion, (Pokemondata$HP) ^ 0.2 is the best transformation since it removes some of outliers and gives the symmetric data distribution.  

#For Salaries in SF 
We are going to reexpress the variable Total.benefits to be symmetric. 
```{r}
SalariesForSF <- read.csv("https://docs.google.com/spreadsheets/d/1tqh_FERElW3C0Z_Rn2NP76SZdsbKsrViqjw_vtQEFww/pub?output=csv")
hist(SalariesForSF$Total.benefits)
symplot(SalariesForSF$Total.benefits)
```

There is slightly left-skewedness in the symplot since the points fall under the red rine. There is one outlier in the tail.

```{r}
letter.values <- lval(SalariesForSF$Total.benefits)
letter.values$mids
plot(letter.values$mids)
```

You can see the left skewness since this sequence is in decreasing. However, in the tail you can see that the value jumps.

```{r}
#take roots
roots <- sqrt(SalariesForSF$Total.benefits)
aplpack::stem.leaf(roots)
root.lv <- lval(roots)
root.lv
plot(root.lv$mids)

#take log
logs <- log(SalariesForSF$Total.benefits) 
aplpack::stem.leaf(logs)
logs.lv <- lval(logs)
logs.lv
plot(logs.lv$mids)

#if take p = −1/2     -1/sqrt(lake$Area) Further investigation 
recroots <- - 1 / sqrt(SalariesForSF$Total.benefits)
aplpack::stem.leaf(recroots)
recroots.lv <- lval(recroots)
recroots.lv
plot(recroots.lv$mids)

hinkley(roots)
hinkley(logs)
hinkley(recroots)

```

Looking at the values of d in Hinkley's quick method, the correct reexpression appears to be grater than p=.5, although the value closest to 0 corresponds to the roots reexpression.

Therefore, we are gaoing to take a new power transforation as p=3
```{r}
new<-(SalariesForSF$Total.benefits) ^ 3
hist(new, main="P = 3")
aplpack::stem.leaf(new)
new.lv <- lval(new)
new.lv
plot(new.lv$mids)
symplot(new)
```

Now the date is reexpressed as roughly symmetric. However, there is one huge outlier so we have to be careful when we go to the further investigation. 

```{r}
#let x_0 = 65, the medain of the raw data
x_0<-68528.65
matched.roots <- x_0 + (sqrt(SalariesForSF$Total.benefits) - sqrt(x_0)) / (.5*x_0 ^ (-.5))
matched.logs <- x_0 + (log10(SalariesForSF$Total.benefits) - log10(x_0)) / (log10(exp(1)) / x_0)

raw <-SalariesForSF$Total.benefits
matched.roots <- mtrans(raw, 0.5)
matched.logs <- mtrans(raw, 0)
boxplot(data.frame(raw, matched.roots, matched.logs))
```

Now,all of data look like symmetric. While raw data has wider data spreads including outlier, matched.logs data has more narrow data spreads. 

In conclusion, this transformation,(SalariesForSF$Total.benefits) ^ 3, gives the roughly symmetric distribution. However, it is not the best of the best because there is one huge outlier that gives the right-skewedness. We have met the data which is difficult to reexpress. 
